The Blue Obelisk five years on

### 2011-03-02 paper.tex
More on OSOS

It is worth noting that Pharao is the first Open Source pharmacophore program (\emph{Is this a reasonable statement?) (although pharmacophore-related functionality has been available in the CDK and RDKit for some time).

The availability of feature-rich and robust cheminformatics toolkits has started to bear fruit in the development of higher-level tools and applications that are built using these toolkits.

Progress to date...
  \subsection*{Remaining challenges}
What's left to do?

>>>>

It is worth noting that Pharao is the first Open Source pharmacophore program (\emph{Is this a reasonable statement?}) (although pharmacophore-related functionality has been available in the CDK and RDKit for some time).

  \subsection*{Remaining challenges}

Accuracy. Very often software work at the 95\% to 98\% level. The variety of chemical structures is such that with a large enough dataset, 'unusual' structures are always found which may be mishandled by software. Given that much of the development of open source software is unfunded, and relies heavily on developer motivation, it is understandable that working on the final N\% of problem structures (which may require substantial work) is not the most exciting of tasks. Putting a bounty system in place may be useful for these situations where a researcher needs a problem to be fixed because it affects a particular dataset of interest.

Performance. There is a famous quote by Knuth that "Premature optimization is root of all evil (or something)". Given that compute time is cheap, ....

Databases?

\section*{Open Standards}
  \subsection*{Progress}

OpenSMILES.

InChI

CML?

  \subsection*{Remaining challenges}

OpenSMARTS.

Structure Normalisation

\section*{Open Data}
  \subsection*{Progress}
  \subsection*{Remaining challenges}
